The error in Bayesian phylogenetic reconstruction when speciation co-occurs

### 2018-12-04 article.tex
Differentiate between number of taxa that are extant (n) and are extant or extinct (N)

\item Here we present our study with the aim of exploring 
when using a more complex MBD tree prior is warranted.
To do so, we simulate phylogenies using the MBD process, with
varying degrees of that process. To be explicit, we define
that degree, $s$, as the number of extinct and extant species created
during a co-occuring speciation event, 
$n_{\mathbb{M}}$, from the total number of extinct and extant species:

\begin{equation}
s = \frac{n_{\mathbb{M}}}{n_{\mathbb{M}} + n_{\mathbb{B}}}
\end{equation}

Here, $n_{\mathbb{B}}$ is the number of extinct and extant species created
during a default single-birth speciation event.

@@

\item Current phylogenetic tools assume that only a single speciation event 
can occur at any given time.
While this assumption is useful to construct a wide variety of successful 
models (e.g \cite{Maddison2007biSSE}, \cite{Valente2015}, 
\cite{etienne2012diversity}, \cite{etienne2014estimating}),
they disallow for environmental changes that trigger speciations 
in multiple clades at a same point in time.

\item In the MBD model, parameters $\lambda$ and $\mu$ correspond, respectively, 
to the common per-species speciation and extinction rates present 
also in the standard BD model. 
Additionally, MBD relies on two additional parameters. 
Parameter $\nu$ is the rate at which an environmental change is triggered.
When such event is triggered, 
all species present in the phylogeny at that moment
have a probability $q$ to speciate at that time, which is 
independent on $\lambda$. 
Polytomies are not allowed in such process 
as each species can speciate only once at the time.

@@

\item To prove our hypothesis we simulate two twin datasets. 
All the simulations are produced in continuous time, 
using the Doob-Gillespie algorithm.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
  \centering 
  \begin{tabular}{r l}
    \hline
                          & Description\\
    \hline
    \hline
    $e$                   & Inference error \\
    $n_{\mathbb{M}}$      & Number of extinct and extant taxa created by a multiple-birth event \\
    $n_{\mathbb{B}}$      & Number of extinct and extant taxa created by a single-birth event \\
    $s$                   & Extent/strength of the multiple birth process \\
    $L_c$                 & MCMC chain length \\
    $L_{sc}$              & MCMC sub-chain length \\
    $\epsilon$            & relative error tolerance in marginal likelihood estimation \\
    \hline
  \end{tabular}
  \caption{
    Overview of the symbols. 
  }
  \label{table:symbols}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
  \centering 
  \begin{tabular}{r l}
    \hline
                          & Value(s) \\
    \hline
    \hline
    $L_c$                 & $10^6$ \\
    $L_{sc}$              & $10^4$ \\
    $\epsilon$            & $10^{-13}$ \\
    \hline
  \end{tabular}
  \caption{
    Overview of the simulation parameters. 
  }
  \label{table:parameters}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

\item Here we present our study with the aim of exploring 
when using a more complex MBD tree prior is warranted.
To do so, we simulate phylogenies using the MBD process, with
varying degrees of that process. To be explicit, we define
that degree, $s$, as the number of extinct and extant species created
during a co-occuring speciation event, 
$N_{\mathbb{M}}$, from the total number of extinct and extant species:

\begin{equation}
s = \frac{N_{\mathbb{M}}}{N_{\mathbb{M}} + N_{\mathbb{B}}}
\end{equation}

Here, $N_{\mathbb{B}}$ is the number of extinct and extant species created
during a default single-birth speciation event.

@@

\item 
\richel{TODO: move to Introduction}
Current phylogenetic tools assume that only a single speciation event 
can occur at any given time.
While this assumption is useful to construct a wide variety of successful 
models (e.g \cite{Maddison2007biSSE}, \cite{Valente2015}, 
\cite{etienne2012diversity}, \cite{etienne2014estimating}),
they disallow for environmental changes that trigger speciations 
in multiple clades at a same point in time.

\item 
\richel{TODO: move to Introduction}
In the MBD model, parameters $\lambda$ and $\mu$ correspond, respectively, 
to the common per-species speciation and extinction rates present 
also in the standard BD model. 
Additionally, MBD relies on two additional parameters. 
Parameter $\nu$ is the rate at which an environmental change is triggered.
When such event is triggered, 
all species present in the phylogeny at that moment
have a probability $q$ to speciate at that time, which is 
independent on $\lambda$. 
Polytomies are not allowed in such process 
as each species can speciate only once at the time.

@@

\item To investigate the effect of $s$ on $e$, we simulate phylogenies
for different values of $s$ spread equally from zero (no multiple-birth
event, thus a BD model) to one (species are created only from multiple-birth
events) and three intermediate values of 0.25, 0.5 and 0.75.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
  \centering 
  \begin{tabular}{r l}
    \hline
                          & Description\\
    \hline
    \hline
    $e$                   & Inference error \\
    $n_{\mathbb{M}}$      & Number of extant taxa created by a multiple-birth event \\
    $n_{\mathbb{B}}$      & Number of extant taxa created by a single-birth event \\
    $N_{\mathbb{M}}$      & Number of extinct and extant taxa created by a multiple-birth event \\
    $N_{\mathbb{B}}$      & Number of extinct and extant taxa created by a single-birth event \\
    $s$                   & Extent/strength of the multiple birth process \\
    $L_c$                 & MCMC chain length \\
    $L_{sc}$              & MCMC sub-chain length \\
    $\epsilon$            & relative error tolerance in marginal likelihood estimation \\
    \hline
  \end{tabular}
  \caption{
    Overview of the symbols. 
  }
  \label{table:symbols}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
  \centering 
  \begin{tabular}{r l}
    \hline
                          & Value(s) \\
    \hline
    \hline
    $s$                   & 0.0, 0.25, 0.5, 0.75, 1.0 \\
    $L_c$                 & $10^6$ \\
    $L_{sc}$              & $10^4$ \\
    $\epsilon$            & $10^{-13}$ \\
    \hline
  \end{tabular}
  \caption{
    Overview of the simulation parameters. 
  }
  \label{table:parameters}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
